---
title: "Meteorites Weekend Homework"
author: "Lesley Duff"
date: "`r format(Sys.Date())`"
output:
  html_document:
    df_print: paged
subtitle: Analysis
---
# Meteorites Analysis

We have data from [NASA](https://www.nasa.gov/) which has information on 
meteorites that have been found up to the year 2013.

```{r}
library(testthat)
```

```{r}
source("cleaning.R") # brings in libraries assertr/janitor/tidyverse

# Read the cleaned data into R.
meteorite_data <- process_meteorite_landings_file("data/meteorite_landings.csv")
```
## Names and years for the 10 largest meteorites

```{r}
# Find the names and years found for the 10 largest meteorites in the data.
# PDA 3.5 Writing function/program to perform statistical analysis
meteorite_data %>%
  slice_max(mass_g, n = 10) %>%
  select(name, year)
```
## Average mass of meteorites

```{r}
# Find the average mass of meteorites that were recorded falling, vs. those
# which were just found.
# PDA 3.5 Writing function/program to perform statistical analysis
meteorite_data %>%
  group_by(fall) %>%
  summarise(avg_mass_g = mean(mass_g, na.rm = TRUE))
```

## Number of meteorites in each year, for every year since 2000.

```{r}
# Find the number of meteorites in each year, for every year since 2000.
# PDA 3.5 Writing function/program to perform statistical analysis
meteorite_data %>%
  filter(year >= 2000) %>%
  group_by(year) %>%
  summarise(number_of_meteorites = n())
```

## Function get_meteorite_count()

```{r}
# Write a function called get_meteorite_count() that takes in two arguments:
#  a dataframe and a year, and returns the phrase "In (input_year), there were
# (number of meteorites that year) meteorites." IF there are 1 or more
# meteorites that year. If there are no meteorites in the year, return the
# phrase "As far as we know, there were no meteorites that year".
# 3.5 Writing function/program to perform statistical analysis
get_meteorite_count <- function(dataframe, input_year) {
  output <- ""

  # Extract number_of_meteorites that year
  number_of_meteorites <- dataframe %>%
    filter(year == input_year) %>%
    summarise(number_of_meteorites = n()) %>%
    pull()

  if (number_of_meteorites >= 1) {
    output <- str_c(
      "In ", input_year, ", there were ", number_of_meteorites,
      " meteorites."
    )
  } else {
    output <- "As far as we know, there were no meteorites that year"
  }

  return(output)
}

# Year with count
count_meteorite <- get_meteorite_count(meteorite_data, 2009)
count_meteorite

# Year with no count
count_meteorite <- get_meteorite_count(meteorite_data, 2020)
count_meteorite

# count_meteorite <- get_meteorite_count(meteorite_data, NA)
# count_meteorite <- get_meteorite_count(meteorite_data, -1000)
```
### Test plans

#### How to test that function is working as expected 

```{r}
# 4.1 Test plans 
test_that("Years with meteorites return count", {
  expect_equal(
    get_meteorite_count(meteorite_data, 2009),
    "In 2009, there were 103 meteorites."
  )
  expect_equal(
    get_meteorite_count(meteorite_data, 2013),
    "In 2013, there were 2 meteorites."
  )
})

test_that("Years with no meteorites return no meteorites message", {
  expect_equal(
    get_meteorite_count(meteorite_data, 1584),
    "As far as we know, there were no meteorites that year"
  )
})
```

How the function could be improved.

1. The simple if statement means negative numbers are treated the same as zero.
Add an extra test explicitly for zero and indicate anything else is a negative 
number and an error.

```{r, error=TRUE}
# 4.1 Test plans 
test_that("Year cannot be negative", {
  expect_error(get_meteorite_count(meteorite_data, -2013))
})
```
2. There is no checking to make sure that year is not in the future 
 e.g. mistyping a year
```{r, error=TRUE}
# 4.1 Test plans 
test_that("Year is not in the future", {
  expect_error(get_meteorite_count(meteorite_data, 20012))
})
```

3. The NASA dataset only goes up to the year 2013. 
It is stated in the notes that the dataset provided is for "meteorites that 
have been found up to the year 2013"
Treat years > 2013 <= current year as errors.
Meteorite counts for years after 2013 have the potential to mislead users as 
the message "As far as we know, there were no meteorites that year" could be 
interpreted as meaning that there were *zero* meteorites when all we can say is
that no data has been made available past 2013.
"Absence of Evidence is not [Evidence of Absence](https://en.wikipedia.org/wiki/Evidence_of_absence)" - Carl Sagan.
N.B. Adding this check could potentially make the code brittle if new data file 
with more recent data comes along. So this test would be reconsidered if we were 
aware of a new dataset

```{r, error=TRUE}
# 4.1 Test plans 
test_that("NASA dataset only goes up to year 2013", {
  expect_error(get_meteorite_count(meteorite_data, 2014))
})
```

### Unit test

Test to check whether non-numeric inputs 
(character and logical) for the year argument return an error.

```{r, error=TRUE}
test_that("non-numeric inputs for the year argument return an error", {
  expect_error(get_meteorite_count(meteorite_data, TRUE))
  expect_error(get_meteorite_count(meteorite_data, "2009"))
})
```

### refactor_get_meteorite_count

```{r}
# 4.6 Performing test and revising code
refactor_get_meteorite_count <- function(dataframe, input_year) {
  # Make sure input_year only takes in numeric input
  if (!is.numeric(input_year)) {
    stop("Non-numeric year 😿")
  }
  # Make sure number is >= 0
  if (input_year < 0) {
    stop("Year cannot be a negative number 😿")
  }
  # Check that year isn't in future
  current_year <- as.numeric(format(Sys.Date(), "%Y"))
  if (input_year > current_year) {
    stop("Year cannot be in the future 😿")
  }
  if (input_year > 2013) {
    stop("NASA data only available up to year 2013 😿")
  }

  output <- ""

  # Extract number_of_meteorites that year
  number_of_meteorites <- dataframe %>%
    filter(year == input_year) %>%
    summarise(number_of_meteorites = n()) %>%
    pull()

  output <- case_when(
    # Return a count if there is a positive number of landing
    number_of_meteorites >= 1 ~
      str_c(
        "In ", input_year, ", there were ", number_of_meteorites,
        " meteorites."
      ),
    # Explicitly test for zero
    number_of_meteorites == 0 ~
      "As far as we know, there were no meteorites that year",
    TRUE ~ "As far as we know, there were no meteorites that year"
  )

  return(output)
}
```

```{r, error=TRUE}
# 4.6 Performing test and revising code
# Test your function to make sure it now gives an error when you input a logical
# or character value as the year.
test_that("Non-numeric input returns an error", {
  expect_error(refactor_get_meteorite_count(meteorite_data, "year"))
  expect_error(refactor_get_meteorite_count(meteorite_data, TRUE))
})

# Check working as expected
test_that("Years with meteorites return count", {
  expect_equal(
    get_meteorite_count(meteorite_data, 2009),
    "In 2009, there were 103 meteorites."
  )
  expect_equal(
    get_meteorite_count(meteorite_data, 2013),
    "In 2013, there were 2 meteorites."
  )
})

test_that("Years with no meteorites return no meteorites message", {
  expect_equal(
    get_meteorite_count(meteorite_data, 1584),
    "As far as we know, there were no meteorites that year"
  )
})

# 4.6 Performing test and revising code
# My own suggested improvements
# Error if negative number supplied
test_that("Year cannot be negative", {
  expect_error(refactor_get_meteorite_count(meteorite_data, -2013))
})

test_that("Year is not in the future", {
  expect_error(refactor_get_meteorite_count(meteorite_data, 2024))
  expect_error(refactor_get_meteorite_count(meteorite_data, 20012))
})

test_that("NASA dataset only goes up to year 2013", {
  expect_error(refactor_get_meteorite_count(meteorite_data, 2014))
})
```
