---
title: "Meteorites Weekend Homework"
subtitle: "Analysis"
output: html_notebook
date: "`r format(Sys.Date())`"
author: "Lesley Duff"
---
# Meteorites Analysis

We have data from [NASA](https://www.nasa.gov/) which has information on 
meteorites that have been found up to the year 2013.

```{r}
source("cleaning.R")

# Read the cleaned data into R.
meteorite_data <- process_meteorite_landings_file("data/meteorite_landings.csv")
```
## Names and years for the 10 largest meteorites

```{r}
# Find the names and years found for the 10 largest meteorites in the data.
meteorite_data %>%
  slice_max(mass_g, n = 10) %>%
  select(name, year)
```
## Average mass of meteorites

```{r}
# Find the average mass of meteorites that were recorded falling, vs. those
# which were just found.
meteorite_data %>%
  group_by(fall) %>%
  summarise(avg_mass_g = mean(mass_g, na.rm = TRUE))
```

## Number of meteorites in each year, for every year since 2000.

```{r}
# Find the number of meteorites in each year, for every year since 2000.
meteorite_data %>%
  filter(year >= 2000) %>%
  group_by(year) %>%
  summarise(number_of_meteorites = n())
```

## Function get_meteorite_count()

```{r}
# Write a function called get_meteorite_count() that takes in two arguments:
#  a dataframe and a year, and returns the phrase "In (input_year), there were
# (number of meteorites that year) meteorites." IF there are 1 or more
# meteorites that year. If there are no meteorites in the year, return the
# phrase "As far as #we know, there were no meteorites that year".
get_meteorite_count <- function(dataframe, input_year) {
  output <- ""
  
  # Extract number_of_meteorites that year
  number_of_meteorites <- dataframe %>%
    filter(year == input_year) %>%
    summarise(number_of_meteorites = n()) %>%
    pull()

  #  number_of_meteorites <- 2
  #  number_of_meteorites <- 0
  if (number_of_meteorites >= 1) {
    output <- str_c(
      "In ", input_year, ", there were ", number_of_meteorites,
      " meteorites."
    )
  } else {
    output <- "As far as we know, there were no meteorites that year"
  }

  return(output)
}

count_meteorite <- get_meteorite_count(meteorite_data, 2009)
#count_meteorite <- get_meteorite_count(meteorite_data, NA)
#count_meteorite <- get_meteorite_count(meteorite_data, -1000)
print(count_meteorite)
```
### test plans

How the function could be improved.

1. We know there are several years in the dataset with NAs, if NA is passed as 
a year the message could be improved to show something sensible when the year is
unknown. Check whether NA passed as year parameter and return new output

2. The simple if statement means negative numbers are treated the same as zero.
Add an extra test explicitly for zero and indicate anything else is a negative 
number and an error. Create a new output if year is numeric but < 0

3. There is no type checking of parameters. Would check if first argument is 
dataframe and year is numeric.
