---
title: "Meteors Continued, Again"
subtitle: Decision Tree
output:
  html_document:
    df_print: paged
---
# Decision Tree
```{r}
library(rpart)
library(rpart.plot)
library(tidyverse)
```

This is a mini homework for people doing the PDA. You are going to continue 
working on your week 2 weekend homework.

Based on the mass and year, create a decision tree model of your meteorites 
data which predicts whether a meteorite was seen falling, or was discovered 
after its impact (fell/found).

First, you want to convert any character column to a factor. Next (and most 
importantly) split your dataset into training and testing datasets. Then 
create a decision tree with rpart and plot it using rpart.plot. Finally, test 
and add your predictions to your data, and create a confusion matrix.

```{r}
# Load clean meteorite data
meteorite_landings <- read_csv("clean_data/meteorite_landings.csv")
```

```{r}
#View(meteorite_landings)

# Meterorites are ordered by year shuffle them
shuffle_index <- sample(1:nrow(meteorite_landings))

# shuffle the data so class order isn't in order - need this for training/testing split later on 
meteorite_landings_shuffled <- meteorite_landings[shuffle_index, ]

# First, you want to convert any character column to a factor. 
meteorite_landings_factored <- meteorite_landings_shuffled %>%
  mutate(across(where(is.character), as.factor))

head(meteorite_landings_factored)

skimr::skim(meteorite_landings_factored)

meteorite_landings_factored <- meteorite_landings_factored %>% 
  # We still have some missing entries for year lets clean them
  na.omit() %>% 
  
  #remove some unneeded columns
  select(-id, -name)

glimpse(meteorite_landings_factored)

```
split your dataset into training and testing datasets.


```{r}
# 80% of our data going into the training set, and the 
# remaining 20% go in our test set.

# get how many rows we have in total to work out the percentage
n_data <- nrow(meteorite_landings_factored)

# create a test sample index
test_index <- sample(1:n_data, size = n_data*0.2)

# create test set
meteorite_landings_test  <- slice(meteorite_landings_factored, test_index)

# create training set
meteorite_landings_train <- slice(meteorite_landings_factored, -test_index)
```


 check that our test and training sets have similar proportions of 
fall. 

```{r}
meteorite_landings_test %>%
 janitor::tabyl(fall)
```

```{r}
meteorite_landings_train %>%
 janitor::tabyl(fall)
```

Build tree model based on training dataset
Then create a decision tree with rpart

```{r}
meteorite_landings_fit <- rpart(
  #?rpart
  # Based on the mass and year
  # mass_g and year,
  formula = fall ~ mass_g + year,  
  data = meteorite_landings_train, 
  method = 'class'
)
#meteorite_landings_fit
```
plot it using rpart.plot

```{r}
rpart.plot(meteorite_landings_fit, 
           yesno = 2, 
           fallen.leaves = TRUE, 
           faclen = 6, 
           digits = 2)
```


