---
title: "Meteors Continued, Again"
subtitle: Decision Tree
output:
  html_document:
    df_print: paged
---
# Decision Tree
```{r}
library(modelr)
library(rpart)
library(rpart.plot)
library(tidyverse)
library(yardstick)
```

This is a mini homework for people doing the PDA. You are going to continue 
working on your week 2 weekend homework.

Based on the mass and year, create a decision tree model of your meteorites 
data which predicts whether a meteorite was seen falling, or was discovered 
after its impact (fell/found).

First, you want to convert any character column to a factor. Next (and most 
importantly) split your dataset into training and testing datasets. Then 
create a decision tree with rpart and plot it using rpart.plot. Finally, test 
and add your predictions to your data, and create a confusion matrix.

```{r}
# Load clean meteorite data
meteorite_landings <- read_csv("clean_data/meteorite_landings.csv")
```

```{r}
# View(meteorite_landings)

# Meterorites are ordered by year shuffle them
shuffle_index <- sample(1:nrow(meteorite_landings))

# shuffle the data so class order isn't in order - need this for training/testing split later on
meteorite_landings_shuffled <- meteorite_landings[shuffle_index, ]

# First, you want to convert any character column to a factor.
meteorite_landings_factored <- meteorite_landings_shuffled %>%
  mutate(across(where(is.character), as.factor))

head(meteorite_landings_factored)

skimr::skim(meteorite_landings_factored)

meteorite_landings_factored <- meteorite_landings_factored %>%
  # We still have some missing entries for year lets clean them
  na.omit() %>%
  # remove some unneeded columns
  select(-id, -name)

glimpse(meteorite_landings_factored)
```
split your dataset into training and testing datasets.


```{r}
# 80% of our data going into the training set, and the
# remaining 20% go in our test set.

# get how many rows we have in total to work out the percentage
n_data <- nrow(meteorite_landings_factored)

# create a test sample index
test_index <- sample(1:n_data, size = n_data * 0.2)

# 4.2 Testing datasets (Week 11)
# create test set
meteorite_landings_test <- slice(meteorite_landings_factored, test_index)

# 4.3 Training datasets (Week 11)
# create training set
meteorite_landings_train <- slice(meteorite_landings_factored, -test_index)
```


 check that our test and training sets have similar proportions of 
fall. 

```{r}
# 4.2 Testing datasets (Week 11)
meteorite_landings_test %>%
  janitor::tabyl(fall)
```

```{r}
# 4.3 Training datasets (Week 11)
meteorite_landings_train %>%
  janitor::tabyl(fall)
```

Build tree model based on training dataset
Then create a decision tree with rpart

```{r}
# 4.3 Training datasets (Week 11)
meteorite_landings_fit <- rpart(
  # ?rpart
  # Based on the mass and year
  # mass_g and year,
  formula = fall ~ mass_g + year,
  data = meteorite_landings_train,
  method = "class"
)
# meteorite_landings_fit
# Do once to save tree
# saveRDS(meteorite_landings_fit, file = "clean_data/meteorite_landings_fit.RDS")
```
plot it using rpart.plot

```{r}
meteorite_landings_fit <- readRDS(file = "clean_data/meteorite_landings_fit.RDS")

# 4.3 Training datasets (Week 11)
rpart.plot(meteorite_landings_fit,
  yesno = 2,
  fallen.leaves = TRUE,
  faclen = 6,
  digits = 2
)
```

Finally, test and add your predictions to your data


```{r}
# 4.2 Testing datasets (Week 11)
# add the predictions
meteorite_landings_test_pred <- meteorite_landings_test %>%
  add_predictions(meteorite_landings_fit, type = "class")

# look at the variables  choose the ones that our decision tree model showed
# as most informative.
# names(meteorite_landings_test)

meteorite_landings_test_pred %>%
  select(year, mass_g, fall, pred)
```

check predictive performance by creating a confusion matrix
```{r}
# 4.2 Testing datasets (Week 11)
conf_mat <- meteorite_landings_test_pred %>%
  conf_mat(truth = fall, estimate = pred)

conf_mat
```
Accuracy

```{r}
accuracy <- meteorite_landings_test_pred %>%
 accuracy(truth = fall, estimate = pred)

accuracy 
```
Sensitivity

```{r}
sensitivity <- meteorite_landings_test_pred %>%
  sensitivity(truth = fall, estimate = pred)

sensitivity
```
Specificity

```{r}
specificity <- meteorite_landings_test_pred %>%
  specificity(truth = fall, estimate = pred)

specificity
```

