---
title: "Penguins analysis part 2"
subtitle: "Correlation - PDA"
author: "Lesley Duff"
date: "`r format(Sys.Date())`"
output: html_document
---
# Correlation
```{r message=FALSE, warning=FALSE}
library(GGally)
library(ggplot2)
library(tidyverse)
```

```{r}
# read in cleaned data
penguins <- read_csv("../clean_data/penguins.csv")
```
## Correlation matrix

Create a correlation matrix using GGally::ggcorr() that shows the correlation between bill length, bill depth, flipper length, and body mass across species. (Hint: you can add correlation coefficient labels to your plot if you look at the ?ggcorr help file)

Write a short summary of your results.

```{r include=FALSE}
# My cleaned data on the previous analysis dropped the other columns
# I need to revisit the cleaning script
# Old clean = species and bill length
# Need to add bill depth, flipper length and body mass

names(penguins)
#[1] "species"           "bill_length_mm"    "bill_depth_mm"    
#[4] "flipper_length_mm" "body_mass_g"  
```

### Correlation between bill length, bill depth, flipper length, and body mass across species

```{r warning=TRUE}
# 4.5 Concept of correlation and correlation matrices
corr_data_penguins <- penguins %>% 
  select(bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g)
#correlation between bill length, bill depth, flipper length, and body mass across species.
#View(penguins)
# ?ggcorr

# Pearson correlation coefficients, using pairwise observations 
# (default method)
# Pearson’s correlation:
# corresponds to the covariance of the two variables normalized (i.e., divided) by the product of their standard deviations.

# By default, the midpoint of the gradient is set at 0
#, which indicates a null correlation

ggcorr(corr_data_penguins, label = TRUE)
```

* Strong positive correlation between flipper length and body mass.  
Close to nearly perfect linear relationship.
The longer the flippers the heavier the penguin
* strong positive correlation between bill length and flipper length and body mass. Longer bills mean longer flippers and heavier penguins
* Bill depth strong negative correlation  with flipper length and body mass, So as bill depth increases, flipper length and body mass tend to decrease
* Bill length and and bill depth relatively weak negative correlation. As bill length increases, bill depth tends to decrease and vice versa, but the relationship is not very strong.


