---
title: "Penguins analysis part 1"
subtitle: "Statistics & Hypothesis testing - PDA"
author: "Lesley Duff"
date: "`r format(Sys.Date())`"
output:
  html_document:
    df_print: paged
---
# Penguins analysis part 1
```{r}
library(e1071)
library(infer)
library(tidyverse)

penguins <- read_csv("../clean_data/penguins.csv")
```
We are interested in how bill length differ between the three species of penguin
(Adelie, Chinstrap & Gentoo) in this dataset.

## 1. Exploratory analysis

```{r}
# How many of each species?
penguins %>%
  ggplot() +
  geom_bar(aes(x = species),
    fill = "steelblue",
    colour = "white",
    stat = "count"
  ) +
  theme_minimal() +
  labs(
    x = "\nPenguin species",
    y = "Count\n",
    title = "Penguin count by species",
    fill = "Species"
  )
```

```{r}
# Calculate mean bill length by species
mean_penguins <- penguins %>%
  select(species, bill_length_mm) %>%
  group_by(species) %>%
  summarise(mean_bill_length_mm = mean(bill_length_mm)) %>%
  arrange(desc(mean_bill_length_mm))

mean_penguins %>%
  ggplot() +
  geom_col(aes(x = species, y = mean_bill_length_mm),
    fill = "steelblue",
    colour = "white"
  ) +
  theme_minimal() +
  labs(
    x = "\nPenguin species",
    y = "Mean bill length (mm)\n",
    title = "Mean penguin bill length by species"
  )
```

```{r}
# 4.1 Exploratory data analysis and targeted data analysis (hypothesis testing)
penguins %>%
  ggplot() +
  geom_histogram(
    aes(
      x = bill_length_mm,
      fill = species
    ),
    col = "white",
    fill = "steelblue",
    binwidth = 1
  ) +
  facet_wrap(~species, ncol = 3) +
  theme_minimal() +
  labs(
    x = "\nPenguin bill length (mm)",
    y = "Count\n",
    title = "Penguin bill length by species"
  )

mean_penguins
```

### Summary
* There are more Adelie penguins than any other species, they have the shortest mean bill length. 
* There are fewest of Chinstrap penguins these have the longest mean bill length.
* Chinstrap and Gentoo appear to have a bimodal distribution

## 2. Summary statistics

Calculate the following summary statistics to describe bill length for each species:

* measures of centrality (mean & median)
* spread (standard deviation)
* skewness


### Bill length summary statistics
```{r}
# 4.4 Summary statistics including measures of centrality, spread and skewness
penguin_summary_stats <- penguins %>%
  group_by(species) %>%
  summarise(
    count = n(),
    range = max(bill_length_mm) - min(bill_length_mm),
    mean = mean(bill_length_mm),
    median = median(bill_length_mm),
    sd = sd(bill_length_mm),
    skewness = skewness(bill_length_mm, type = 1)
  )
penguin_summary_stats
```


* Adelie = fairly symmetrical, slight right skew
* Chinstrap = fairly symmetrical, with slight left skew,
highest standard deviation, observations are more spread out	
* Gentoo = moderately right skewed


## 3. Hypothesis test & probability

Perform a suitable hypothesis test to check whether the mean bill length of Gentoo is significantly longer than the mean bill length of Adelie.

Mean bill length Gentoo = 47.50488

Mean bill length Adelie = 38.79139	


$H_0$: $\mu_{\textrm{bill_length_mm}}(Gentoo)$−$\mu_{\textrm{bill_length_mm}}(Adelie)$=0

$H_a$: $\mu_{\textrm{bill_length_mm}}(Gentoo)$−$\mu_{\textrm{bill_length_mm}}(Adelie)$>0

$\alpha = 0.05$

```{r}
# 4.1 Exploratory data analysis and targeted data analysis (hypothesis testing)
# 4.3 Probability and probability distributions

observed_stat <- penguins %>%
  filter(species %in% c("Gentoo", "Adelie")) %>%
  specify(bill_length_mm ~ species) %>%
  calculate(stat = "diff in means", order = c("Gentoo", "Adelie"))

observed_stat

null_distribution <- penguins %>%
filter(species %in% c("Gentoo", "Adelie")) %>%
  specify(bill_length_mm ~ species) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 10000, type = "permute") %>%
  calculate(stat = "diff in means", order = c("Gentoo", "Adelie"))

null_distribution %>%
  visualise(bins = 30)+
  shade_p_value(obs_stat = observed_stat$stat, 
                direction = "right")
```

How likely is it to see a result as extreme as your observed result, if the null hypothesis is true?

```{r}
p_value <- null_distribution %>%
  get_p_value(obs_stat = observed_stat$stat, 
              direction = "right")
p_value
```
p <  $\alpha$

p of 0 observed difference is unlikely to be due to chance.

We reject $H_0$ as the p value is less than $\alpha$
 
There is strong evidence to conclude that the mean bill length of Gentoo is significantly longer than the mean bill length of Adelie. These are independent
samples.
