---
title: "Penguins analysis part 1"
subtitle: "Statistics & Hypothesis testing - PDA"
author: "Lesley Duff"
date: "`r format(Sys.Date())`"
output:
  html_document:
    df_print: paged
---
# Penguins analysis part 1
```{r}
library(tidyverse)

penguins <- read_csv("../clean_data/penguins.csv")
```
We are interested in how bill length differ between the three species of penguin
(Adelie, Chinstrap & Gentoo) in this dataset.

## 1. Exploratory analysis

```{r}
# How many of each species?
penguins %>%
  ggplot() +
  geom_bar(aes(x = species),
    fill = "steelblue",
    colour = "white",
    stat = "count"
  ) +
  theme_minimal() +
  labs(
    x = "\nPenguin species",
    y = "Count\n",
    title = "Penguin count by species",
    fill = "Species"
  )
```

```{r}
# Calculate mean bill length by species
mean_penguins <- penguins %>%
  select(species, bill_length_mm) %>%
  group_by(species) %>%
  summarise(mean_bill_length_mm = mean(bill_length_mm)) %>%
  arrange(desc(mean_bill_length_mm))

mean_penguins %>%
  ggplot() +
  geom_col(aes(x = species, y = mean_bill_length_mm),
    fill = "steelblue",
    colour = "white"
  ) +  
  theme_minimal() +
  labs(
    x = "\nPenguin species",
    y = "Mean bill length (mm)\n",
    title = "Mean penguin bill length by species"
  )
```

```{r}
penguins %>%
  ggplot() +
  geom_histogram(
    aes(
      x = bill_length_mm,
      fill = species
    ),
    col = "white",
    fill = "steelblue",
    binwidth = 1
  ) +
  facet_wrap(~species, ncol = 3) +
  theme_minimal() +
  labs(
    x = "\nPenguin bill length (mm)",
    y = "Count\n",
    title = "Penguin bill length by species"
  )

mean_penguins
```

### Summary
* There are more Adelie penguins than any other species, they have the shortest mean bill length. 
* There are fewest of Chinstrap penguins these have the longest mean bill length.

## 2. Summary statistics

## 3. Hypothesis test & probability
