---
title: 'PDA Data Transformation Task'
output:
  html_document:
    df_print: paged
author: "Lesley Duff"
date: "`r format(Sys.Date())`"
---
# Data Transformation Task
```{r}
library(tidyverse)
```

## 1
Load in the transaction data and create a column date from the day, month and year columns.
```{r}
toy_transactions <- read_csv("raw_data/toy_transactions.csv")

toy_transactions_with_date <- toy_transactions %>%
  mutate(date = as.Date(make_datetime(year, month, day)))
toy_transactions_with_date
```
## 2
Load in the toys data and:

1. Convert the weight column to numeric.
2. Split the extra information from the product into a new column, so you have two columns: product_name and product_title. Make sure you don’t have any extra whitespace in either column.

```{r}
toys <- read_csv("raw_data/toys.csv")
toys_clean <- toys %>%
  # Extract weight as a number
  mutate(weight = parse_number(weight)) %>%
  # Split into new column
  separate(product, c("product_name", "product_title"), " - ") %>%
  # remove any leading/trailing space
  mutate(
    product_name = str_trim(product_name),
    product_title = str_trim(product_title)
  )

toys_clean
```

## 3
Load in the dataset on quality labels and:

1. Remove the unnecessary information for each descriptor
2. Replace the categories so ‘Awesome’ and ‘Very Awesome’ become ‘Good’ and 
‘Very Good’. Do the same thing for ‘Awful’ replacing it with ‘Bad’.
```{r}
quality <- read_csv("raw_data/quality.csv")
# View(quality)

quality_clean <- quality %>%
  # Remove the unnecessary information for each descriptor
  mutate(description = str_replace_all(description, "Quality: ", "")) %>%
  # Replace the categories so ‘Awesome’ and ‘Very Awesome’ become ‘Good’
  # and ‘Very Good’. Do the same thing for ‘Awful’ replacing it with ‘Bad
  mutate(
    description = case_when(
      description == "Awesome" ~ "Good",
      description == "Very Awesome" ~ "Very Good",
      description == "Awful" ~ "Bad",
      
      # I'm assuming 'Very Awful' has to be treated like 'Very Awesome' but if
      # meant literally *only* 'Awful' I would comment the following line out
      description == "Very Awful" ~ "Very Bad",
      TRUE ~ description
    )
  )

quality_clean
```


## 4
## 5
## 6
## 7
## 8
## 9
## 10
## 11
