---
title: "Preliminary Analysis"
output:
  html_document:
    df_print: paged
---
# CitiBike Preliminary Analysis
This is a temporary area for examining the CitiBike dataset in terms of quality and range of data available.

This information is used to create the data cleaning script.

```{r}
library(tidyverse)
library(tsibbledata)

nyc_bikes_raw <- nyc_bikes

#?nyc_bikes
#View(nyc_bikes_raw)

# Time series of class tsibble for start_time/stop_time

#str(nyc_bikes_raw)
# gives "interval")=Error in `format()`:

# 4268 row of 12 variables

```
* start_time: The time and date when the trip was started.
* stop_time: The time and date when the trip was ended.
* start_station: A unique identifier for the starting bike station.
* start_lat: The latitude of the starting bike station.
* start_long: The longitude of the starting bike station.
* end_station: A unique identifier for the destination bike station.
* end_lat: The latitutde of the destination bike station.
* end_long: The longitude of the destination bike station.
* type: The type of trip. A "Customer" has purchased either a 24-hour or 3-day pass, and a "Subscriber" has purchased
birth_year The bike rider’s year of birth.
*gender: The gender of the bike rider.

```{r}
# explore variables
head(nyc_bikes_raw)
#?POSIXct

glimpse(nyc_bikes_raw)

# check for NAs
skimr::skim(nyc_bikes_raw)
```

## Bikes
```{r}
nyc_bikes_raw %>% 
  count(bike_id, sort = TRUE)

```
There are 10 different bikes in this sample.

## Bike stations
```{r}
nyc_bikes_raw %>% 
  count(start_station, sort = TRUE)

nyc_bikes_raw %>% 
  count(end_station, sort = TRUE)

```

## User Types
```{r}
#‘Customer’ purchased a short-term 24-hour or 3-day pass
#‘Subscriber’ is a long-term Annual Member

nyc_bikes_raw %>% 
    count(type, sort = TRUE)

# Vast majority 3953 are already subscribers - good!
# Only 315 are using the short term options - opportunties to grow 'casual' use?
```


## Demographics

```{r}
nyc_bikes_raw %>% 
  count(gender, sort = TRUE)
# Male/Female/Unknown

# Percent unknown?
269/4268 * 100

# Spread of ages
nyc_bikes_raw %>% 
  count(birth_year, sort = FALSE)

nyc_bikes_raw %>% 
  distinct(birth_year)

# There are a couple of birth_years in 1880s. Take a look at records < 1900
nyc_bikes_pre_1900 <- nyc_bikes_raw %>%
  filter (birth_year < 1900)
nyc_bikes_pre_1900
# Only two trips, possible user error choosing year? Likely to be 1980s?

nyc_bikes_raw %>% 
  ggplot(aes(birth_year)) +
    geom_histogram()

```

## Journeys

```{r}
# We have start and end times. How long was the hire in duration?
nyc_bikes_hire_duration <- nyc_bikes_raw %>% 
  mutate(hire_duration =  as.duration(stop_time - start_time),
         hire_duration_minutes = round(as.numeric(hire_duration, "minutes")),
         .after = stop_time)
   # mutate(time_duration  = as.duration(today() - date_added))

##View(nyc_bikes_trip_duration)
nyc_bikes_trip_duration %>% 
  ggplot(aes(trip_duration_minutes)) +
  geom_histogram()

## Some entries over multiple days

nyc_bikes_trip_duration %>% 
  filter(trip_duration_minutes < 24 * 60) %>% 
  ggplot(aes(trip_duration_minutes)) +
  geom_histogram(binwidth = 10)
# Vast majority of hires < 200 minutes in total


```

