---
title: "Preliminary Analysis"
output:
  html_document:
    df_print: paged
---
# CitiBike Preliminary Analysis
This is a temporary area for examining the CitiBike dataset in terms of quality and range of data available.

This information is used to create the data cleaning script.

```{r}
library(leaflet) # For spatial maps
library(tidyverse)
library(tsibbledata)

nyc_bikes_raw <- nyc_bikes

# ?nyc_bikes
# View(nyc_bikes_raw)

# Time series of class tsibble for start_time/stop_time

# str(nyc_bikes_raw)
# gives "interval")=Error in `format()`:

# 4268 row of 12 variables
```
* start_time: The time and date when the trip was started.
* stop_time: The time and date when the trip was ended.
* start_station: A unique identifier for the starting bike station.
* start_lat: The latitude of the starting bike station.
* start_long: The longitude of the starting bike station.
* end_station: A unique identifier for the destination bike station.
* end_lat: The latitutde of the destination bike station.
* end_long: The longitude of the destination bike station.
* type: The type of trip. A "Customer" has purchased either a 24-hour or 3-day pass, and a "Subscriber" has purchased
birth_year The bike rider’s year of birth.
*gender: The gender of the bike rider.

```{r}
# explore variables
head(nyc_bikes_raw)
# ?POSIXct

glimpse(nyc_bikes_raw)

# check for NAs
skimr::skim(nyc_bikes_raw)
```

## Bikes
```{r}
nyc_bikes_raw %>%
  count(bike_id, sort = TRUE)
```
There are 10 different bikes in this sample.

## Bike stations
```{r}
nyc_bikes_raw %>%
  count(start_station, sort = TRUE)
# 52 different starting stations

nyc_bikes_raw %>%
  count(end_station, sort = TRUE)
# 55 different end stations

nyc_bikes_raw %>%
  count(start_station, end_station, sort = TRUE) %>%
  # ?count
  slice_head(n = 5)
# top 5 start/end combinations
# 1023 different combinations
# Most popular journey is between station 3203 and 3186
```
### Locations of Bike Stations
```{r}
nyc_bikes_station_locations <- nyc_bikes_raw %>%
  select(start_station, start_lat, start_long) %>%
  group_by(start_station,start_lat, start_long) %>%
  summarise(number_of_hires = n()) %>%
  ungroup()
nyc_bikes_station_locations

nyc_bikes_station_start_map <- leaflet(nyc_bikes_station_locations) %>%
  addTiles() %>% # use the default base map which is OpenStreetMap tiles
  #?addMarkers

  addCircleMarkers(
    lng = ~ start_long,
    lat = ~ start_lat,
    radius = ~ number_of_hires/20,
    popup = ~paste("<strong>Number of Hires:</strong>", number_of_hires,
      "<br /><strong>Station Number:</strong>", 
                   start_station),
    label = ~paste("Number of Hires:",number_of_hires)
  ) 

nyc_bikes_station_start_map
```




## User Types
```{r}
# ‘Customer’ purchased a short-term 24-hour or 3-day pass
# ‘Subscriber’ is a long-term Annual Member

nyc_bikes_raw %>%
  count(type, sort = TRUE)

# Vast majority 3953 are already subscribers - good!
# Only 315 are using the short term options - opportunties to grow 'casual' use?
```


## Demographics

```{r}
nyc_bikes_raw %>%
  count(gender, sort = TRUE)
# Male/Female/Unknown

# Percent unknown?
269 / 4268 * 100

# Spread of ages
nyc_bikes_raw %>%
  count(birth_year, sort = FALSE)

nyc_bikes_raw %>%
  distinct(birth_year)

# There are a couple of birth_years in 1880s. Take a look at records < 1900
nyc_bikes_pre_1900 <- nyc_bikes_raw %>%
  filter(birth_year < 1900)
nyc_bikes_pre_1900
# Only two trips, possible user error choosing year? Likely to be 1980s?

nyc_bikes_raw %>%
  ggplot(aes(birth_year)) +
  geom_histogram()
```

## Journeys

```{r}
# We have start and end times. How long was the hire in duration?
nyc_bikes_hire_duration <- nyc_bikes_raw %>%
  mutate(
    hire_duration = as.duration(stop_time - start_time),
    hire_duration_minutes = round(as.numeric(hire_duration, "minutes")),
    .after = stop_time
  )
# mutate(time_duration  = as.duration(today() - date_added))

#nyc_bikes_hire_duration_categories <- nyc_bikes_hire_duration %>% 
  #select(hire_duration) %>% 
  #mutate(trip_duration_type = 
           # < 10
           # 10 - 19
           # 20 - 29
           # >= 30
   #        )
#nyc_bikes_hire_duration_categories

## View(nyc_bikes_trip_duration)
nyc_bikes_hire_duration %>%
  ggplot(aes(x = hire_duration_minutes)) +
  geom_histogram()

## Some entries over multiple days

nyc_bikes_hire_duration %>%
  filter(hire_duration_minutes < 24 * 60) %>%
  ggplot(aes(hire_duration_minutes)) +
  geom_histogram(binwidth = 10)
# Vast majority of hires < 200 minutes in total
```
