---
title: "Preliminary Analysis"
output:
  html_document:
    df_print: paged
---
# Preliminary Analysis
This is a temporary area for examining the CitiBike dataset.

```{r}
library(tidyverse)
library(tsibbledata)

nyc_bikes_raw <- nyc_bikes

#?nyc_bikes
#View(nyc_bikes_raw)

# Time series of class tsibble for start_time/stop_time

#str(nyc_bikes_raw)
# gives "interval")=Error in `format()`:

# 4268 row of 12 variables

```
* start_time: The time and date when the trip was started.
* stop_time: The time and date when the trip was ended.
* start_station: A unique identifier for the starting bike station.
* start_lat: The latitude of the starting bike station.
* start_long: The longitude of the starting bike station.
* end_station: A unique identifier for the destination bike station.
* end_lat: The latitutde of the destination bike station.
* end_long: The longitude of the destination bike station.
* type: The type of trip. A "Customer" has purchased either a 24-hour or 3-day pass, and a "Subscriber" has purchased
birth_year The bike rider’s year of birth.
*gender: The gender of the bike rider.

```{r}
# explore variables
head(nyc_bikes_raw)
#?POSIXct

glimpse(nyc_bikes_raw)

# check for NAs
skimr::skim(nyc_bikes_raw)
```

## Bikes
```{r}
nyc_bikes_raw %>% 
  count(bike_id, sort = TRUE)

```
There are 10 different bikes in this sample.

## Bike stations
```{r}
nyc_bikes_raw %>% 
  count(start_station, sort = TRUE)

nyc_bikes_raw %>% 
  count(end_station, sort = TRUE)

```



## Demographics

```{r}
nyc_bikes_raw %>% 
  count(gender, sort = TRUE)
# Male/Female/Unknown

# Percent unknown?
269/4268 * 100

# Spread of ages
nyc_bikes_raw %>% 
  count(birth_year, sort = FALSE)

nyc_bikes_raw %>% 
  distinct(birth_year)

# There are a couple of birth_years in 1880s. Take a look at records < 1900
nyc_bikes_pre_1900 <- nyc_bikes_raw %>%
  filter (birth_year < 1900)
nyc_bikes_pre_1900
# Only two trips, possible user error choosing year? Likely to be 1980s?

nyc_bikes_raw %>% 
  ggplot(aes(birth_year)) +
    geom_histogram()

```


